Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-04-07 Cost_manuscript.Rmd
edited methods section, adding figures

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1,2 & Joseph T. McGuire^1,2"
csl: apa.csl
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: default #reference_docx: Style_reference_draft.docx
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#bibliography: mPFC_paper.bib
---

@@

  # do p < 0.001 or round depending on the value
  adapted_p <- ifelse(pval < 0.001, "< 0.001", paste("=", round(pval, digits = 2)))

@@

1.  Department of Psychological and Brain Sciences, Boston University, Boston, USA

2.  Center for Systems Neuroscience, Boston University, Boston, USA

@@

	We excluded participant datasets based on four preregistered criteria: 1) Consent: If they withdrew their participation; 2) Inattentiveness: a catch trial was placed at the end of each experimental block, asking participants to press a key within 3 seconds (time requirement based on pilot study response times). A participant that failed two or more of these checks was excluded and replaced. 3) Improbable choice behavior: The task was structured so that one reward amount must always be accepted. A participant who quit every trial in at least one block was assumed not to have followed or understood task instructions, or to have disengaged from the task altogether. 4) Performance: In the cognitive effort condition, participants were forced to travel if they made 2 mistakes in a trial (see task procedures below). Any participant with more than 30% forced travels was excluded.

@@

  All experimental tasks were implemented using PsychoPy 2.X (CITE) on a Macbook Pro laptop. In this task, participants foraged for monetary rewards in an environment in which each trial required either physical effort or cognitive effort for a set period of time (the “handling time”), or an equivalent unfilled delay. Their goal was to maximize their gains within a fixed amount of time. Participants exerted physical effort by maintaining grip on a handheld dynamometer (XXXX, BIOPAC Systems, United States) using their dominant hand. Gripping requirements were calibrated at 20% of maximum voluntary contraction (MVC, acquired at the beginning of the session). Cognitive effort entailed switching among Stroop, dot motion coherence, and flanker tasks. In Stroop, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. word red painted in red) or incongruent (e.g. word red painted in blue). Participants had to select the color of the font, not the word displayed (i.e. they had to suppress their tendency to read the word). For motion coherence, 100 solid white dots moved on the screen. A fraction of these dots moved cohesively to the left or right (coherence could be either 30% or 40%, uniformly sampled), while the rest moved in random directions. Participants had to respond with the direction of the cohesive set of dots. REVISE. In flanker, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction from its neighbors. These tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time. Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive ones of each kind.

  Each participant was assigned to one of these types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling. If participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), such that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude experience-dependent learning.

@@

  Foraging theory posits that accepting a delayed reward depends on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens and Krebs, 1977). In this study, the richness of the environment was manipulated by the length of the handling and travel times (i.e. shorter travel times produced richer environments). Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

**INSERT**

  where R, H, and P are the reward, handling time, and acceptance probability of trial i, respectively, and T is the travel times. This gave us the total reward per second attainable in each block as a function of the lowest amount accepted (i.e. acceptance threshold). Figure X shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. Beyond comparing among cost types, this approach allowed us to test if certain demands boosted the value of offers (i.e. over-accepting low offers).

@@

  First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a logistic regression to each participant’s data to predict trial-wise acceptances, using handling time and reward amount as predictors. The resulting beta coefficients for handling time and reward were pooled across all participants, and we performed a one-sample rank-sum test on each set of coefficients to examine whether they were significantly positive or negative (compared to no effect of 0). We then performed an extension of the logistic regression by adding an autoregressive covariate containing the number of consecutive quits prior to a given trial. This examined the possibility that participant choices were governed by recent quitting history rather than the experimental parameters. Coefficients significantly different from 0 denoted that a participant relied on recent quitting history.

	Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion accepted (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost above). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR). **MAYBE CHANGE TO COST3 APPROACH?**

  Next, we compared preferences among cost conditions. We first performed a one-way ANOVA on the proportion of trials accepted using group as a factor. We then compared the proportion completed among all 4 groups using non-parametric permutation contrasts (6 tests). On each of 5000 permutation iterations, the group assignment was randomly shuffled without replacement, and the difference in mean acceptance rates across iterations created a putative null distribution. The unpermuted group mean difference was then evaluated against this permuted distribution. The same approach was used for total earnings. This gave us an initial glimpse on the potential differences in cost among conditions. In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random effect. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. As preregistered, we anticipated significant main effects (coefficients different than zero) of handling time, reward, and cost condition. In addition, we hypothesized that the probability of accepting would be significantly great in either effort condition compared to wait and easy groups. We then examined whether this a priori model outperformed both simpler and more complex models. We used both Akaike’s Information Criterion (AIC) and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in the following order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 5 would have the lowest AIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Finally, we modeled the subjective opportunity cost in each condition. We started by addressing the prediction that participants would display consistent choice patterns throughout the session (i.e. that the reward amounts they accepted in a given timing condition would be similar throughout the experiment). We computed each participant’s total proportion of acceptances pre- and post-midpoint for every block, and performed paired permutations (5000) on the mean proportion of acceptances. We then confirmed that the relative costs observed with from the whole experimental session were still present on each half of the experiment. The winning model from the mixed-effects logistic comparison was applied to each half separately. In each case the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost above). OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

 **INSERT**

  Both gamma and the temperature parameter of the softmax function were estimated independently for each subject, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. Finally, the OC estimates for each group were compared using an ANOVA with condition as a factor. This let us determine which cost type produced the highest discounting. **ADD ADDITIONAL MODELS FROM LITERATURE**

@@

  We collected data from a new sample of 48 eligible participants  (39 Female, median age = 21, range = 18 - 36). Sample size was once again determined by means of power analysis (repeated measures ANOVA), using a significance level of 0.05, power of 0.8, a desired effect size of f = 0.5, and four factors (one for each condition). The resulting sample size was 45, which we increased to 48 in order to balance the potential order of blocks. We excluded participants using the same criteria as in the between-subject experiment.

@@

  The original between-subjects task was modified so that participants foraged for cost types in addition to rewards. Every participant experienced all forms of cost (delay, physical effort, and cognitive effort). There were two block types, in which delay trials were interspersed with either physical-effort or cognitive-effort trials. This setup prevented participants from having to rapidly switch between response modalities across trials (i.e. keyboard press and handgrip). Block combinations (i.e. physical/wait and cognitive/wait) were experienced three times by each participant in 7-minute-long, interleaved blocks, thus matching the experiment’s length to the between-subject version. Half of the participants experienced a block sequence that started with the physical block. Timing parameters were matched to the middle condition from Experiment 1 (i.e. 10 s handling and 6 s travel), allowing a balance between acquiring enough observations per trial type and being able to compare results across experiments. Participants were informed that this timing combination would be constant prior to beginning the experiment. Cost type combinations were disclosed at the beginning of each block, and reward/cost offers displayed before each trial begins (e.g. “8 cents for physical effort”). Unlike the previous experiment, participants could make their choice during the offer window (instead of during the handling time). Participants trained in each type of demand until they reached the same criteria as in the between-subject experiment.

@@

  The analytical pipeline for the within-subject study was mostly preserved from the between-subject experiment. We first ran tests to determine whether decision makers integrated reward information. We performed a logistic regression per participant to predict trial-wise acceptances based on reward amounts. The resulting beta coefficients were pooled across all participants, and a one-sample rank-sum test was used to examine whether they were significantly positive or negative (compared to 0). Next, we explored the effects of choice history by adding an autoregressive covariate containing the number of consecutive recent quits. We then confirmed that the under- and over-acceptance were detrimental to total earnings by performing the same quadratic linear model explained above. To determine the optimality of the group’s decisions, we examined whether each cost type produced a bias to over or under-accept at 4 and 8 cents (assuming that 20 cents were always be accepted, given our design). The reward-maximizing strategy was to always reject 4 cents and accept 8 cents, yielding a combined optimal proportion of acceptances of 50% (see Operationalization of Costs for details). We performed a two-sided one-sample chi-squared test of proportions against the null probability of 0.5 for each type of cost. Therefore, a significant difference indicated that participants either over or under accepted rewards.

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and subject ID as a random effect (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in the following order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

@@

  **adapt once we decide on the model candidates**To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

@@

``` {r BTW: 16.1.3. Earnings, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# calculate the proportions and earnings per subject, while keeping group info
temp <- dataBtw %>%
  filter(rawChoice < 2) %>%
  group_by(SubjID, Cost) %>%
  summarise(Earnings = sum(Offer[Choice == 1] / 100),
            pComplete = mean(Choice)) %>%
  mutate(pComplete2 = pComplete^2) %>%
  ungroup()

# linear fit of earnings x prop accept
lmEarn <- lm(Earnings ~ pComplete + pComplete2, data = temp)

# plot
# (earnFitplot <- ggplot(data = temp, aes(pComplete, Earnings)) + 
#          scale_color_manual(values = colsBtw) + 
#          geom_point(aes(fill = Cost), pch = 21, color = "black", size = 3, show.legend = T) + 
#          stat_smooth(aes(y = Earnings), method = "lm", formula = y ~ x + I(x^2), color = "grey30") +
#          labs(x = "Proportion Completed", y = "Total Earnings (dollars)") +
#          scale_fill_manual(values = colsBtw) +
#          theme(panel.grid.major = element_blank(), 
#            panel.grid.minor = element_blank(), 
#            panel.background = element_blank(), 
#            axis.line = element_line(colour = "black"),
#            text = element_text(size = 16)))

@@

  In this experiment, groups faced with different behavioral costs (physical effort, cognitive effort, or delay) were tasked with deciding their preferred strategy to maximize rewards in fully-disclosed foraging environments of variable richness. Environmental richness was dictated by the time it took to obtain a reward (the handling time) and the time between trials (the travel time). We hypothesized that participants would approximate reward-maximizing behavior by preferring higher rewards and shorter delays, but that changing the cost required to obtain rewards would uniformly increase or decrease this choice pattern. Specifically, we predicted that participants confronted with cognitive demands would be more likely to accept trials than those faced with physical effort and delay, regardless of handling time and reward amount.

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey & Joseph T. McGuire"
csl: apa.csl
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: default #reference_docx: Style_reference_draft.docx
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#bibliography: XXXX.bib
---

@@

  # do p < 0.001 or round depending on the value
  # maybe adapt with case_when(), so there are three levels: observed, < 0.01, and < 0.001
  adapted_p <- ifelse(pval < 0.001, "< 0.001", paste("=", round(pval, digits = 3)))

@@

Department of Psychological and Brain Sciences & Center for Systems Neuroscience, Boston University, Boston, USA

@@

We excluded participant datasets based on four preregistered criteria: 1) Consent: If they withdrew their participation; 2) Inattentiveness: a catch trial was placed at the end of each experimental block, asking participants to press a key within 3 seconds (time requirement based on pilot study response times). A participant that failed two or more of these checks was excluded and replaced. 3) Improbable choice behavior: The task was structured so that one reward amount must always be accepted. A participant who quit every trial in at least one block was assumed not to have followed or understood task instructions, or to have disengaged from the task altogether. 4) Performance: Participants were forced to travel if they made 2 mistakes in a cognitive effort trial (see task procedures below), or if they gripped below threshold during physical trials. Any participant with more than 30% forced travels was excluded.

@@

  All experimental tasks were implemented using PsychoPy 2 (v1.85.1, CITE) on a Macbook Pro laptop. In this task, participants foraged for monetary rewards in an environment in which each trial required either physical effort or cognitive effort for a set period of time (the “handling time”), or an equivalent unfilled delay (Figure 1). Their goal was to maximize their gains within a fixed amount of time. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling.

  Each participant was assigned to one of three types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. Participants exerted physical effort by maintaining grip on a handheld dynamometer (XXXX, BIOPAC Systems, United States) using their dominant hand. Gripping requirements were calibrated at 20% of maximum voluntary contraction (MVC, acquired at the beginning of the session). Cognitive effort entailed switching among Stroop, dot motion coherence, and flanker tasks. In Stroop, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. word red painted in red) or incongruent (e.g. word red painted in blue). Participants had to select the color of the font, not the word displayed (i.e. they had to suppress their tendency to read the word). For motion coherence, 100 solid white dots moved on the screen. A fraction of these dots moved cohesively to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the cohesive set of dots. In flanker, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction from its neighbors. These tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time. Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive tasks of each kind. If participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), with the constraint that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude experience-dependent learning.

```{r Trial plot (Fig. 1), out.width = "80%", fig.align = "center", fig.cap= "General foraging trial structure. On each trial, participants were offered to earn money (4, 8, or 20 cents) by sustaining effort or waiting during the handling time (2, 10, or 14 s). The end of a trial was followed by a travel time (handling and travel times always added up to 16 s). Participants could skip unfavorable trials and immediately start traveling to a potentially better offer. In the between-subject experiment, cost was fixed per participant and handling time varied per block. In the within-subject version, handling time was fixed at 10 s, but a combination of effort and delay trials changed per block. Possible costs, handling times, and rewards were fully disclosed in order to avoid experience-dependent learning."}

include_graphics("./Images/FIG1_short.png")
```

@@

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens and Krebs, 1977). In this study, the richness of the environment was manipulated by the length of the handling and travel times (i.e. shorter travel times produced richer environments). Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

$\dfrac {\sum p_i R_i} {\sum p_i H_i + 3T},p \in \{0, 1\}$

  where *R*, *H*, and *p* are the reward, handling time, and acceptance probability of offer i, respectively, and T is the travel time. This gave us the total reward per second attainable in each block as a function of the lowest amount accepted (i.e. acceptance threshold). Figure 2 shows possible earnings for each choice strategy, as well as the lowest amount participants should accept in order to maximize their rewards (circled dots). For example, a participant in a 10 s handling block should accept 8-cent and 20-cent rewards (and reject 4-cent rewards) to maximize their reward rate. Note that accepting every offer was often detrimental to participant earnings. Significant deviations from the optimal strategy can be understood as reflecting changes in subjective OC. Beyond comparing among cost types, this approach allowed us to test if certain demands boosted the value of offers (i.e. over-accepting low offers).

```{r Optimal plot (Fig. 2), out.width = "80%", fig.align = "center", fig.cap= "Possible earnings per second for each acceptance threshold (i.e. the smallest amount accepted) for each handling time. Circles denote the reward-maximizing threshold for each block, which is described in the table. Experiments shared the 10 s handling block, in which it was optimal to skip all 4 cent offers."}

include_graphics("./Images/FIG2.png")
```

@@

  All analyses were performed in R 3.5.2 (CITE). First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a logistic regression to each participant’s data to predict trial-wise acceptances, using handling time and reward amount as predictors. The resulting beta coefficients for handling time and reward were pooled across all participants, and we performed a one-sample rank-sum test on each set of coefficients to examine whether they were significantly positive or negative (compared to no effect of 0). We then performed an extension of the logistic regression by adding an autoregressive covariate containing the number of consecutive quits prior to a given trial. This examined the possibility that participant choices were governed by recent quitting history rather than the experimental parameters. Coefficients significantly different from 0 denoted that a participant relied on recent quitting history.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion accepted (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR).

  Next, we compared preferences among cost conditions. We first performed a one-way ANOVA on the proportion of trials accepted using group as a factor. We then compared the proportion completed among all 4 groups using non-parametric permutation contrasts (6 tests). On each of 5000 permutation iterations, the group assignment was randomly shuffled without replacement, and the difference in mean acceptance rates across iterations created a putative null distribution. The unpermuted group mean difference was then evaluated against this permuted distribution. The same approach was used for total earnings. This gave us an initial glimpse on the potential differences in cost among conditions.

  In order to further look at the effect of handling, offer, and cost type, we computed the probability of accepting a trial with a mixed-effects logistic regression using the lme4 package (CITE). Based on the task structure and our main question, our a priori model of interest included cost condition, handling time, and reward amount as fixed main effects, and participant ID as a random intercept. Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition. We ran three versions of the model with different reference conditions, in order to test all pairwise differences among the four cost conditions. (**exponentiate coefficients to make them more interpretable?**) We then examined whether this a priori model outperformed both simpler and more complex models. We used both Akaike’s Information Criterion (AIC) and Bayesian Information Criterion (BIC) to determine the model that minimized the negative log-likelihood while penalizing the addition of parameters. The regression with each combination of predictors was fitted in the following order: 1) intercept only; 2) condition only; 3) handling time only; 4) offer amount only; 5) handling time and offer amount (added to the preregistered comparison); 6) condition, handling time, and reward main effects (a priori model from above); 7) adding a handling-by-reward interaction; and 8) all possible two-way interactions. We predicted that model 6 (the a priori setup) would have the lowest AIC and BIC. Nested models with similar AIC were statistically compared using an analysis-of-deviance. The significance test was computed as the probability of the reduction in deviance, based on a chi-square distribution with degrees of freedom equal to the difference in the number of parameters between models.

  Finally, we modeled the subjective opportunity cost in each condition. We started by examining whether participants were confident and consistent in their choices. We visually examined survival curves indicating at what point during the handling time participants quit a trial (censored points included trial completions and forced travels), in addition to quitting time distributions. Next, we addressed the prediction that participants would display consistent choice patterns throughout the session (i.e. that the reward amounts they accepted in a given timing condition would be similar throughout the experiment). We computed each participant’s total proportion of acceptances pre- and post-midpoint for every block, and performed paired permutations (5000 iterations) to compare their mean proportion of acceptances. We then confirmed that the relative costs observed with from the whole experimental session were still present on each half of the experiment. The winning model from the mixed-effects logistic comparison was applied to each half separately. In each case, the reference cost category was rotated in order to assess pairwise differences among all costs. We evaluated the consistency of the relative costs by visually comparing the patterns of significant differences between experimental halves.

  Once we established choice stability, we proceeded to estimate the subjective OC of each type of demand (see Operationalization of Cost). OC was computed as the product of a free parameter (gamma) and the handling time. Gamma can be interpreted as the subjective value of each second spent in an environment. If a trial's reward rate (i.e. offer amount divided by the handling time) fell below gamma, a participant was better off finding more profitable alternatives. In other words, individuals with higher gammas required higher rewards in order to accept a given offer. We used a softmax function to model each participant’s probability of completing a trial based on the difference between the delayed reward’s magnitude and the estimated OC for each cost type

$P(accept)_i = \dfrac {1} {1 + e^{-\beta(R_i - \gamma H_i)}}$

  Both gamma ($\gamma$) and the inverse temperature parameter ($\beta$) of the softmax function were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group. We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (0.74, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we determined which cost type produced the highest discounting by comparing gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group. **ADD ADDITIONAL MODELS FROM LITERATURE**

@@

  We collected data from a new sample of 48 eligible participants  (39 Female, median age = 21, range = 18 - 36). Sample size was once again determined by means of a power analysis (repeated measures ANOVA), using a significance level of 0.05, power of 0.8, a desired effect size of f = 0.5, and four factors (one for each condition). The resulting sample size was 45, which we increased to 48 in order to balance the potential order of blocks. We excluded participants using the same criteria as in the between-subject experiment.

@@

  The original between-subjects task was modified so that participants foraged for cost types in addition to rewards. Every participant experienced all forms of cost (delay, physical effort, and cognitive effort). There were two block types, in which delay trials were interspersed with either physical-effort or cognitive-effort trials. This setup prevented participants from having to rapidly switch between response modalities across trials (i.e. keyboard press and handgrip). Block combinations (i.e. physical/wait and cognitive/wait) were experienced three times by each participant in 7-minute-long, interleaved blocks, thus matching the experiment’s length to the between-subject version. Half of the participants experienced a block sequence that started with the physical block. Timing parameters were matched to the middle condition from Experiment 1 (i.e. 10 s handling and 6 s travel), allowing a balance between acquiring enough observations per trial type and being able to compare results across experiments. Participants were informed that this timing combination would be constant prior to beginning the experiment. Cost type combinations were disclosed at the beginning of each block, and reward/cost offers displayed before each trial begins (e.g. “8 cents for physical effort”). Unlike the previous experiment, participants could express their decision to quit during the offer window (instead of during the handling time). Participants trained in each type of demand until they reached the same criteria as in the between-subject experiment.

@@

  The analytical pipeline was mostly preserved from the between-subject experiment. We first ran tests to determine whether decision makers integrated reward information. We performed a logistic regression per participant to predict trial-wise acceptances based on reward amounts. The resulting beta coefficients were pooled across all participants, and a one-sample rank-sum test was used to examine whether they were significantly positive or negative (compared to 0). Next, we explored the effects of choice history by adding an autoregressive covariate containing the number of consecutive recent quits. We then confirmed that the under- and over-acceptance were detrimental to total earnings by performing the same quadratic linear model explained above. To determine the optimality of the group’s decisions, we examined whether each cost type produced a bias to over or under-accept at 4 and 8 cents (assuming that 20 cents were always be accepted, given our design). The reward-maximizing strategy was to always reject 4 cents and accept 8 cents, yielding a combined optimal proportion of acceptances of 50% (see Operationalization of Costs for details). We performed a two-sided one-sample chi-squared test of proportions against the null probability of 0.5 for each type of cost. Therefore, a significant difference indicated that participants either over or under accepted rewards.

  To compare acceptance rates among cost conditions, we first computed the probability of accepting a trial with a mixed-effects logistic regression. The model included cost condition and reward amount as fixed main effects, and participant ID as a random intercept (handling time is not modeled, as participants experienced a single time combination). Cost condition was modeled with three categorical terms, with the fourth condition as the reference condition (two effort conditions, and two delay conditions corresponding to each effort type). We ran three versions of the model with different reference conditions, in order to test all relevant pairwise differences among the four cost conditions. We compared the a priori model to alternative parameterizations using AIC and BIC, varying model complexity in the following order: 1) intercept only; 2) cost only; 3) reward only; 4) cost and reward main effects (a priori model); and 5) adding a two-way interaction. We predicted that model 4 would have the last considerable decrease in the negative log-likelihood. Models with similar AIC were formally compared using the analysis-of-deviance approach from the between-subjects experiment.

@@

  **adapt once we decide on the model candidates** To estimate the subjective opportunity cost of each condition, we used a logistic function to model each participant’s probability of accepting a trial based on the difference between the delayed reward’s magnitude and the estimated opportunity cost (OC) for each cost type. As before, OC was computed as the product of a free parameter (gamma) and the handling time. Both gamma and the temperature parameter of the logistic function were estimated at the subject level, independently for each subject and cost type. We cross-validated each subject’s OC value using the first four blocks (two cognitive, two physical) for estimation, and the last two blocks (one of each effort type) choices for testing. The estimates were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each cost type. The OC estimates for each group were compared using a repeated measures ANOVA with cost as a factor. We performed pairwise paired permutations as post-hoc tests.

@@

``` {r BTW: 16.1.3. Earnings, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
## calculate the proportions and earnings per subject, while keeping group info
temp <- dataBtw %>%
  filter(rawChoice < 2) %>%
  group_by(SubjID, Cost) %>%
  summarise(Earnings = sum(Offer[Choice == 1] / 100),
            pComplete = mean(Choice)) %>%
  mutate(pComplete2 = pComplete^2) %>%
  ungroup()

## linear fit of earnings x prop accept
lmEarn <- lm(Earnings ~ pComplete + pComplete2, data = temp)

## plot
earnFitplot <- ggplot(data = temp, aes(pComplete, Earnings)) +
         scale_color_manual(values = colsBtw) +
         geom_point(aes(fill = Cost), pch = 21, color = "black", size = 3, show.legend = F) +
         stat_smooth(aes(y = Earnings), method = "lm", formula = y ~ x + I(x^2), color = "grey30") +
         labs(x = "Proportion Completed", y = "Total Earnings (dollars)") +
         scale_fill_manual(values = colsBtw) +
         theme(panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           panel.background = element_blank(),
           axis.line = element_line(colour = "black"),
           text = element_text(size = 12))

@@

  In this experiment, groups faced with different behavioral costs (physical effort, cognitive effort, or delay) were tasked with deciding their preferred strategy to maximize rewards in fully-disclosed foraging environments of variable richness. Environmental richness was dictated by the time it took to obtain a reward (the handling time) and the time between trials (the travel time). We hypothesized that participants would approximate reward-maximizing behavior by preferring higher rewards and shorter delays, but that changing the cost required to obtain rewards would uniformly increase or decrease this choice pattern. Specifically, we predicted that participants confronted with effortful demands would be more likely to accept trials than those faced with passive delay, regardless of handling time and reward amount.
